Clone and graft:\\ Testing scientific applications as they are built

### 2015-01-22 turcksin-spring-mechanics/paper.tex
correct shell commands in paper

So how do we achieve this? The general idea is that we save the test cases we
have previously used to convince ourselves that a new part of the program was
correct. To make this work, we have to have three things:
\begin{enumerate}
\item The input file for a previously verified testcase.
\item The output when running the code on this input file; having verified the
  feature, we know or at least believe that this output is correct.
\item A way to run a new version of the code on the output file, to compare
  its current output against the stored output, and to raise a flag if they
  are not the same.
\end{enumerate}
The first two are really not very complicated -- create a directory
\texttt{tests/} in your project and create two files in it, say
\texttt{testcase-1.json} and the corresponding output
\texttt{testcase-1.output}. When it comes time to test at a later time, step 3
above can be as easy as running commands such as
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  python masses-and-springs.py testcase-1.json > testcase-1.output.new
  diff testcase-1.output testcase-1.output.new
\end{lstlisting}
You would repeat all of this for \texttt{testcase-2.json}, of course.
If the \texttt{diff} program produces no output, then the two files are the
same and your program computed the same answer on this input file as it did
back when you created the \texttt{testcase-1.output} file. If there is a
difference, then you know that you just broke one of the testcases. If you
purposefully changed it -- for example because you changed the output format,
or because you fixed behavior you know was erroneous and have verified that it
is now correct, then copy the new output over the new one. If you did not
purposefully change what the program computes, then you know it's time to break out the debugger.

@@

How do we test this new functionality? First, we should check whether the old input files still
work:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  python masses-and-springs.py testcase-1.json > testcase-1.output.new
  diff testcase-1.output testcase-1.output.new
  python masses-and-springs.py testcase-2.json > testcase-2.output.new
  diff testcase-2.output testcase-2.output.new
\end{lstlisting}
With the code that accompanies this paper, \texttt{diff} reports no difference
between the old and the new output files, so this is reassuring.

@@

The solution to these problems that we have ultimately come to use
is a program called \texttt{numdiff}.%
\footnote{See \url{http://www.nongnu.org/numdiff/}.}
\texttt{numdiff} works essentially like
\texttt{diff} but it recognizes if the text it compares in two files is
actually a number. If two numbers differ by less than some relative or
absolute tolerance, then they are considered equal even if the characters by
which they are represented in an output file are different. For example, you
would replace the \texttt{diff} command used previously by the
following to ensure that numbers $a$ and $b$ compare as equal if $|a-b|\le
10^{-6}$ or $|a-b|\le 10^{-8} \min\{|a|,|b|\}$:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  numdiff -a 1e-6 -r 1e-8 testcase-1.output testcase-1.output.new
\end{lstlisting}
Using this trick, numbers that are written into the output file
\texttt{testcase-1.output.new} on one machine with slight differences to the
ones originally saved in \texttt{testcase-1.output} will still be considered
the same, and the comparison will succeed.

>>>>

So how do we achieve this? The general idea is that we save the test cases we
have previously used to convince ourselves that a new part of the program was
correct. To make this work, we have to have three things:
\begin{enumerate}
\item The input file for a previously verified testcase.
\item The output when running the code on this input file; having verified the
  feature, we know or at least believe that this output is correct.
\item A way to run a new version of the code on the output file, to compare
  its current output against the stored output, and to raise a flag if they
  are not the same.
\end{enumerate}
The first two are really not very complicated -- create a directory
\texttt{tests/} in your project and create two files in it, say
\texttt{testcase-1.json} and the corresponding output
\texttt{testcase-1.reference}. When it comes time to test at a later time, step 3
above can be as easy as running commands such as
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  python spring.py testcase-1.json > testcase-1.output
  diff testcase-1.output testcase-1.reference
\end{lstlisting}
You would repeat all of this for \texttt{testcase-2.json}, of course.
If the \texttt{diff} program produces no output, then the two files are the
same and your program computed the same answer on this input file as it did
back when you created the \texttt{testcase-1.reference} file. If there is a
difference, then you know that you just broke one of the testcases. If you
purposefully changed it -- for example because you changed the output format,
or because you fixed behavior you know was erroneous and have verified that it
is now correct, then copy the new output over the new one. If you did not
purposefully change what the program computes, then you know it's time to break out the debugger.

@@

How do we test this new functionality? First, we should check whether the old input files still
work:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  python spring.py testcase-1.json > testcase-1.output
  diff testcase-1.output testcase-1.reference
  python spring.py testcase-2.json > testcase-2.output
  diff testcase-2.output testcase-2.reference
\end{lstlisting}
With the code that accompanies this paper, \texttt{diff} reports no difference
between the old and the new output files, so this is reassuring.

@@

The solution to these problems that we have ultimately come to use
is a program called \texttt{numdiff}.%
\footnote{See \url{http://www.nongnu.org/numdiff/}.}
\texttt{numdiff} works essentially like
\texttt{diff} but it recognizes if the text it compares in two files is
actually a number. If two numbers differ by less than some relative or
absolute tolerance, then they are considered equal even if the characters by
which they are represented in an output file are different. For example, you
would replace the \texttt{diff} command used previously by the
following to ensure that numbers $a$ and $b$ compare as equal if $|a-b|\le
10^{-6}$ or $|a-b|\le 10^{-8} \min\{|a|,|b|\}$:
\begin{lstlisting}[frame=single,basicstyle=\footnotesize,numbers=left,language=csh]
  numdiff -a 1e-6 -r 1e-8 testcase-1.output testcase-1.reference
\end{lstlisting}
Using this trick, numbers that are written into the output file
\texttt{testcase-1.output} on one machine with slight differences to the
ones originally saved in \texttt{testcase-1.reference} will still be considered
the same, and the comparison will succeed.
